---
author: "O. Gimenez"
date: "August 2016"
output: html_document
---
Network analysis of Star Wars: The Attack of the Clones (episode 2) in R
==================================================

This is a free adaptation of two (very) clever analyses made by others:

* [The Star Wars Social Network by Evelina Gabasov](http://evelinag.com/blog/2015/12-15-star-wars-social-network) in which program F# was mostly used to analyse the Star wars social networks 

* [Analyzing networks of characters in 'Love Actually' by David Robinson](http://varianceexplained.org/r/love-actually-network/) in which R was used to analyse the links between the characters of the movie Love Actually.

The aim here is to try and reproduce Evelina's analysis using R only, using David's contribution plus several tweaks I found here and there on the internet. The R code and data are available on my [GitHub page](https://github.com/oliviergimenez). The results is published via [RPub](http://rpubs.com/ogimenez/200849), have a look in particular to the dynamic representation of the network towards the end.

_Disclaimer_: The original blog posts are awesome and full of relevant details, check them out! My objective here was to teach myself how to manipulate data using trendy R packages and do some network analyses. Some comments below have been copied and pasted from these blogs, the credits entirely go to the authors Evelina and David. Last but not least, my code comes with mistakes probably. 

# Read and format data

First, read in data. I found the movie script in doc format [here](theforce.net/timetales/ep2se.doc), which I converted in txt format for convenience. Then, apply various treatments to have the data ready for analysis. I use the old school way for modifying the original dataframe. [Piping](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf) would have made the code more readable, but I do not feel confident with this approach yet.

```{r, echo=TRUE, message=FALSE, warning=FALSE}

# load convenient packages
library(dplyr)
library(stringr)
library(tidyr)

# read file line by line 
raw <- readLines("attack-of-the-clones.txt")

# create data frame
lines <- data_frame(raw = raw) 

# get rid of leading and trailing white spaces
# http://stackoverflow.com/questions/2261079/how-to-trim-leading-and-trailing-whitespace-in-r
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
lines= mutate(lines,raw=trim(raw))

# get rid of the empty lines
lines2 <- filter(lines, raw != "")

# detect scenes: begin by EXT. or INT.
lines3 <-  mutate(lines2, is_scene = str_detect(raw, "T."),scene = cumsum(is_scene)) 

# drop lines that start with EXT. or INT.
lines4 <- filter(lines3,!is_scene)

# distinguish characters from what they say
lines5 <- separate(lines4, raw, c("speaker", "dialogue"), sep = ":", fill = "left",extra='drop')

# read in aliases (from Evelina's post)
aliases <- read.table('aliases.csv',sep=',',header=T,colClasses = "character")
aliases$Alias
aliases$Name

# assign unique name to characters
# http://stackoverflow.com/questions/28593265/is-there-a-function-like-switch-which-works-inside-of-dplyrmutate
multipleReplace <- function(x, what, by) {
  stopifnot(length(what)==length(by))               
  ind <- match(x, what)
  ifelse(is.na(ind),x,by[ind])
}
lines6 <- mutate(lines5,speaker=multipleReplace(speaker,what=aliases$Alias,by=aliases$Name))

# read in actual names (from Evelina's post)
actual.names <- read.csv('characters.csv',header=F,colClasses = "character")
actual.names <- c(as.matrix(actual.names))
# filter out non-characters
lines7 <- filter(lines6,speaker %in% actual.names)

# group by scene
lines8 <- group_by(lines7, scene, line = cumsum(!is.na(speaker))) 

lines9 <- summarize(lines8, speaker = speaker[1], dialogue = str_c(dialogue, collapse = " "))

# Count the lines-per-scene-per-character
# Turn the result into a binary speaker-by-scene matrix
by_speaker_scene <- count(lines9, scene, speaker)
by_speaker_scene

library(reshape2)
speaker_scene_matrix <-acast(by_speaker_scene , speaker ~ scene, fun.aggregate = length)
dim(speaker_scene_matrix)
```

# Analyses

## Hierarchical clustering

```{r}
norm <- speaker_scene_matrix / rowSums(speaker_scene_matrix)
h <- hclust(dist(norm, method = "manhattan"))
plot(h)
```

## Timeline

Use tree to give an ordering that puts similar characters close together
```{r}
ordering <- h$labels[h$order]
ordering
```

This ordering can be used to make other graphs more informative. For instance, we can visualize a timeline of all scenes:
```{r}
scenes <-  filter(by_speaker_scene, n() > 1) # scenes with > 1 character
scenes2 <- ungroup(scenes)
scenes3 <- mutate(scenes2, scene = as.numeric(factor(scene)),
           character = factor(speaker, levels = ordering))
library(ggplot2)
ggplot(scenes3, aes(scene, character)) +
    geom_point() +
    geom_path(aes(group = scene))
```

Create a cooccurence matrix (see [here](http://stackoverflow.com/questions/13281303/creating-co-occurrence-matrix)) containing how many times two characters share scenes
```{r}
cooccur <- speaker_scene_matrix %*% t(speaker_scene_matrix)
heatmap(cooccur)
```

## Social network analyses

### Graphical representation of the network

Here the nodes represent characters in the movies. The characters are connected by a link if they both speak in the same scene. And the more the characters speak together, the thicker the link between them.  
 
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(igraph)
g <- graph.adjacency(cooccur, weighted = TRUE, mode = "undirected", diag = FALSE)
plot(g, edge.width = E(g)$weight)
```

Compute standard network features, degree and betweeness

```{r}
degree(g)
betweenness(g)
```

Get a nicer representation of the network. See [here](http://tagteam.harvard.edu/hub_feeds/1981/feed_items/1388531), for formating from igraph to d3Network.

```{r}
library(d3Network)
library(networkD3)
sg <- simplify(g)
df <- get.edgelist(g, names=TRUE)
df <- as.data.frame(df)
colnames(df) <- c('source', 'target')
df$value <- rep(1, nrow(df))
# get communities
fc <- fastgreedy.community(g)
com <- membership(fc)
node.info <- data.frame(name=names(com), group=as.vector(com))
links <- data.frame(source=match(df$source, node.info$name)-1,target=match(df$target, node.info$name)-1,value=df$value)

forceNetwork(Links = links, Nodes = node.info,Source = "source", Target = "target",Value = "value", NodeID = "name",Group = "group", opacity = 1, opacityNoHover=1)
```

Here the nodes represent characters in the movies. The characters are connected by a link if they both speak in the same scene. The colors are for groups obtained by some algorithms
